\Large{All in the family:\\Anna and Lisa Hahner's finishing times in\\the 2016 women's Olympic marathon}

### 2017-02-03 paper/paper.tex
feb 3 edits

Figure \ref{fig:scatter} contains two plots, both of which describe
how finishing times from the Rio Olympic women's marathon varied as a
function of athletes' personal best times and their Rio half marathon
split times.  The points in the plots are colored by twin/triplet
status, and both plots contain second-order polynomial regression
lines.  We treat an athlete's prior personal best marathon time as a
measure of the athlete's underlying marathon talent, and we interpret
an athlete's half marathon split as a measure of the athlete's skill
at marathon running on August 14, 2016, the date of the women's
marathon in Rio.

\begin{figure}[!ht]
  \caption{Olympic results and measures of marathon ability}
  \label{fig:scatter}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=\textwidth, keepaspectratio]{scatter_plot.pdf}
    \caption{Finishing and personal best times}
    \label{fig:45degreeplot}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=\textwidth, keepaspectratio]{scatter_plot_half.pdf}
    \caption{Finishing and half marathon split times}
    \label{fig:45degreeplot_half}
  \end{subfigure}
\end{figure}

Figure \ref{fig:45degreeplot_half} includes a solid 45-degree line as
the axes in the figure correspond to full marathons; in addition, the
figure's rug marks denote personal best times of runners who DNFed.
Given the paucity of points (five of them) below the pictured
45-degree line, the vast majority of Olympic marathoners ran slower in
Rio compared to their personal bests.  Relative to personal best
times, the Hahner twins were definitely on the slow side, but a number
of runners had greater differences between their Olympic times and
their personal bests than the Hahners.  These runners are denoted with
squares in Figure \ref{fig:45degreeplot}, and there are 13 such
symbols, highlighting approximately 10\% of the racers who completed
the marathon.  The figure shows moreover that there were two women who
had personal best times slightly faster than the Hahner's and yet
finished after the German twins.  Although Figure
\ref{fig:45degreeplot} suggests that the Hahner twins were slower than
one would have expected given their previous best marathon times, it
is not consistent with the accusation that they dramatically slowed
down in the Rio marathon.

@@

\begin{figure}[!ht]
 \caption{Differences in personal best times and finishing times}
 \label{fig:diffdiffscatter}
 \centering
 \begin{subfigure}{.5\textwidth}
 \includegraphics[width=\textwidth, keepaspectratio]{diff_in_diff_scatter_plot.pdf}
 \end{subfigure}
\end{figure}

@@

Consider first the Hahner twins.  They are effectively located on the
horizontal axis because their difference in finishing times is one
second.  However, there are many points about the Hahner's orange dot,
and this shows that, conditional on an approximate 115 second
difference in personal best times, most marathoners did not have close
finishing times like the Hahnes.  Some pairs of runners with around
115 second personal best differences had finishing time differences of
1000 seconds, i.e., in excess of 15 minutes.  Of course the points in
Figure \ref{fig:diffdiffscatter} are not independent, but they provide
a sense of the dispersion in finishing time differences that one can
expect conditional on differences in personal best times.

@@

\begin{figure}[!ht]
    \centering
    \includegraphics[keepaspectratio]{studentized_residuals.pdf}
    \caption{Studentized residuals}
    \label{fig:studentizedresiduals}
  \end{figure}

>>>>

Figure \ref{fig:scatter} and Figure \ref{fig:scatterhalf} contain two plots, both of which describe
how finishing times from the Rio Olympic women's marathon varied as a
function of athletes' personal best times and and as function of their Rio half marathon
split times.  The points in the plots are colored by twin/triplet
status, and both plots contain second-order polynomial regression
lines.  We treat an athlete's prior personal best marathon time as a
measure of the athlete's underlying marathon talent (seen in Figure \ref{fig:scatter}), and we interpret
an athlete's half marathon split as a measure of the athlete's skill
at marathon running on August 14, 2016, the date of the women's
marathon in Rio (seen in Figure \ref{fig:scatterhalf}).

\begin{figure}[!ht]
  \caption{Olympic results as a function of ability measured by personal best}
  \label{fig:scatter}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=\textwidth, keepaspectratio]{scatter_plot.pdf}
    \caption{} 
    \label{fig:45degreeplot}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=\textwidth, keepaspectratio]{studentized_residuals.pdf}
    \caption{}
    \label{fig:studentizedresiduals}
  \end{subfigure}
\end{figure}

\begin{figure}[!ht]
  \caption{Olympic results as a function of ability measured by half marathon split}
  \label{fig:scatterhalf}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=\textwidth, keepaspectratio]{scatter_plot_half.pdf}
    \caption{}
    \label{fig:45degreeplot_half}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=\textwidth, keepaspectratio]{studentized_residuals_half.pdf}
    \caption{}
    \label{fig:studentizedresidualshalf}
  \end{subfigure}
\end{figure}

Figure \ref{fig:45degreeplot} includes a solid 45-degree line as
the axes in the figure correspond to full marathons; in addition, the
figure's rug marks denote personal best times of runners who DNFed.
Given the paucity of points (five of them) below the pictured
45-degree line, the vast majority of Olympic marathoners ran slower in
Rio compared to their personal bests.  Relative to personal best
times, the Hahner twins were definitely on the slow side, but a number
of runners had greater differences between their Olympic times and
their personal bests than the Hahners.  These runners are denoted with
squares in Figure \ref{fig:45degreeplot}, and there are 13 such
symbols, highlighting approximately 10\% of the racers who completed
the marathon.  The figure shows moreover that there were two women who
had personal best times slightly faster than the Hahner's and yet
finished after the German twins.  Although Figure
\ref{fig:45degreeplot} suggests that the Hahner twins were slower than
one would have expected given their previous best marathon times, it
is not consistent with the accusation that they dramatically slowed
down in the Rio marathon.

@@

\begin{figure}[!ht]
 \caption{Differences in personal best times and finishing times}
 \centering
 \begin{subfigure}{.49\textwidth}
 \includegraphics[width=\textwidth, keepaspectratio]{diff_in_diff_scatter_plot.pdf}
  \label{fig:diffdiffscatter}
 \end{subfigure}
  \begin{subfigure}{.49\textwidth}
 \includegraphics[width=\textwidth, keepaspectratio]{diff_in_diff_scatter_plot_half.pdf}
  \label{fig:diffdiffscatterhalf}
 \end{subfigure}
\end{figure}

@@

Consider first the Hahner twins.  They are effectively located on the
horizontal axis because their difference in finishing times is one
second.  However, there are many points about the Hahner's orange dot,
and this shows that, conditional on an approximate 115 second
difference in personal best times, most marathoners did not have close
finishing times like the Hahners.  Some pairs of runners with around
115 second personal best differences had finishing time differences of
1000 seconds, i.e., in excess of 15 minutes.  Of course the points in
Figure \ref{fig:diffdiffscatter} are not independent, but they provide
a sense of the dispersion in finishing time differences that one can
expect conditional on differences in personal best times.

@@
